---
title: "Advanced Filters"
---

This example shows an alternative way of defining dimension or metric filters by using functions from the `ganalytics` package. The current development version of `googleAnalyticsR` supports filters defined with `ganalytics`. `ganalytics` provides functions that makes it easy to define filters using natural R language operators. As of the most recent update to this site, this requires a "dev" (development) version of two packages(both packages exist on CRAN, but they require the dev versions to use the examples here):

* `googleAnalyticsR`
* `ganalytics`

This actually means this example has the incidental educational value of showing how to use packages from GitHub! You can read up on the purpose of `ganalytics` -- developed by [Johann de Boer](https://twitter.com/johannux) -- [on GitHub](https://github.com/jdeboer/ganalytics). But, in short, it's intended to "support R users in defining reporting queries using natural R expressions instead of being concerned about API technical intricacies like query syntax, character code escaping and API limitations." 

# Setup/Config

This example requires development versions of the `googleAnalyticsR` (>=0.5.0.9000) and `ganalytics` (>=0.10.4.9000) R packages available on GitHub, so the setup code below is a bit different (it has some additional code for loading a couple of packages from GitHub).

```{r setup, child='z_setup_ganalytics.Rmd'}
```

# Pull the Data

In this example, we'll define the following filters:
* Device category is desktop or tablet - a dimension filter using an OR condition.
* New visitors using either a desktop or tablet device - a dimension filter involving both an AND and an OR condition.
* At least one goal completion or transaction - a metric filter using an OR condition.

The above list of filters will be defined using `ganalytics` expressions as follows:

```{r define-filters}
# Device category is desktop or tablet - a dimension filter using an OR condition.
desktop_or_tablet <- Expr(~deviceCategory == "desktop") | Expr(~deviceCategory == "tablet")

# New visitors using either a desktop or tablet device - a dimension filter involving both an AND and an OR condition.
new_desktop_and_tablet_visitors <- Expr(~userType == "new") & desktop_or_tablet

# At least one goal completion or transaction - a metric filter using an OR condition.
at_least_one_conversion <- Expr(~goalCompletionsAll > 0) | Expr(~transactions > 0)
```

We can now use `googleAnalyticsR` to pull the data. Chances are, the table below will _only_ show a desktop row. That's because the site we use for this doesn't have a ton of traffic, and it rarely gets tablet conversions! But, trust us, this would return multiple rows on a site with more traffic and conversions.

```{r pull-data, message=FALSE, warning=FALSE, results='hide'}

# Pull the data
results <- google_analytics(
  viewId = view_id,
  date_range = c(start_date, end_date),
  metrics = c("users", "sessions", "goalCompletionsAll", "transactions"),
  dimensions = c("deviceCategory", "userType"),
  dim_filters = new_desktop_and_tablet_visitors,
  met_filters = at_least_one_conversion
)

# Display the results
results
```

```{r display-pulled-data,echo=FALSE, message=FALSE, warning=FALSE}
kable(results)
```

