---
title: "Waterfall Chart / Aggregated Churn"
---

Credit to [Jamarius Taylor](https://www.linkedin.com/in/jamarius-taylor/) for thinking up this one. The idea is that visitors that are new have the opportunity to return to the site (just like customers who buy once have an opportunity to buy more). Obviously, not every site can reasonably expect all visitors to return, and, even those that do will have differing timeframes within which they would _expect_ visitors to return.

Nonetheless, this seems to be an illustrative example. We're going to use a waterfall chart for the visualization. The [waterfalls package](https://cran.r-project.org/web/packages/waterfalls/index.html) offers one option for doing this, but that doesn't exactly seem to be a mainstream package, so this example simply relies on plain ol' `ggplot2` and the various machinations required therein.

# Website Churn

If a website has _zero_ churn, then we would expect each week's users to be the number of users from the _previous week plus the number of _new_ users from the _current_ week. This assumes that active users return at least once per week (which is fair, if, clearly, an ideal that is unachievable):

$$[\#\ of\ Users\ Current\ Week] = [\#\ of\ Users\ Previous\ Week] + [\#\ of\ New\ Users\ Current\ Week]$$

Therefore, one way to calculate the _churn_ is to take that theoretical ideal and subtract the actual number of users from it. That would be the number of non-returning users:

$$[Churned\ Visitors] = [\#\ of\ Users\ Previous\ Week] + [\#\ of\ New\ Users\ Current\ Week] - [\#\ of\ Users\ Current\ Week]$$

The churn _rate_ would then be:

$$[Churn\ Rate] = \frac{[Churned\ Visitors]}{[\#\ of\ Users\ Previous\ Week]}$$

We're not actually going to calculate _Churn Rate_ in this example...but we could.

# Setup/Config

```{r child='z_setup_8_weeks.Rmd'}
```

# Pull the Data

This is pretty straightforward. We just need to pull users and new users by week (this could also be done by week):

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}

# Pull the data. See ?google_analytics_4() for additional parameters.
ga_data <- google_analytics(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = c("users", "newUsers"),
                              dimensions = c("week","year"))

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Munging

## Start by Ordering the Data

This just ensures the data is easy enough to look at and interpret.

```{r munging, message=FALSE, warning=FALSE, results='hide'}

# Using dplyr, sort by year and week.
ga_data <- ga_data %>%
  arrange(year, week)

# Take a quick look at the result.
head(ga_data)
  
```

```{r munge-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

## Calculate the Churned Visitors

```{r churn-calc, message=FALSE, warning=FALSE, results='hide'}

# Calculate the churn and add some date levels
totals_and_churn_df <- ga_data %>%
  mutate(previous_week_users = lag(users, 1, default = 0)) %>%         # Get previous week users
  mutate(churned_users = previous_week_users + newUsers - users) %>%   # Calculate churned users
  mutate(date = paste(year,week, sep="-")) %>%                         # Calculate a "week" as YYYY-WW
  mutate(date = factor(date, levels = date)) %>%                       # Make date a factor
  select(date, users, newUsers, previous_week_users, churned_users)

# NA the churned users and new users for the first row
totals_and_churn_df$churned_users[1] <- NA
totals_and_churn_df$newUsers[1] <- NA

# Display the results
head(totals_and_churn_df)
```

```{r churn-calc-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(totals_and_churn_df))
```

## Prep the Data to Be a Ready for Plotting

Unfortunately, waterfall charts are kind of tricky to do in `ggplot2`. But, we can use 'geom_rect()' to draw this, which means we need to define the `min` and `max` for `x` and `y` for each box on the chart. We'll go ahead and introduce some redundancy in the plotting for the sake of clarity by defining three types of bars that, for each week, will appear in this order:

* *newUsers* -- these are the new users for each week; we'll plot these as green
* *churned_users* -- these are the churned users each week; we'll plot these as red
* *users* -- these are the actual users for each week; we'll plot these as gray

The `xmin` and `xmax` for each of these will differ: *newUsers* will be offset to the left of the week, and *users* will be offset to the right. Yeesh. This is confusing. Ultimately, this is what the data looks like:

```{r waterfall_data, message=FALSE, warning=FALSE, results='hide'}
# Add a few intermediate values that we'll use after we convert to a long forma
waterfall_df <- totals_and_churn_df %>% 
  mutate(week_max = users + churned_users,
         week_current_users = users)

# Convert to long format
waterfall_df <- waterfall_df %>% 
  gather(key = "user_type", value = "number_of_users", 
         -date, -previous_week_users, -week_max, -week_current_users,
         na.rm = TRUE)

# Function to calculate xmin, xmax, ymin, and ymax when passed a row from the
# data set.
calc_rect_coords <- function(date, previous_week_users, week_max,
                             week_current_users, user_type, number_of_users){
  
  x_offset <-  0.25   # Base offset for x for rects
  bar_half_width <- 0.1  # Define the bar width

  if(user_type == "users"){
    xmin <- as.numeric(date) + x_offset - bar_half_width - 0.02  # Make it a little wider
    xmax <- as.numeric(date) + x_offset + bar_half_width + 0.02
    ymin <- 0
    ymax <- number_of_users
    user_type = "Users"
  } else if(user_type == "churned_users"){
    xmin <- as.numeric(date) - bar_half_width
    xmax <- as.numeric(date) + bar_half_width
    ymin <- week_current_users
    ymax <- week_current_users + number_of_users
    user_type = "Churned Users"
  } else 
    # New Users
    {
    xmin <- as.numeric(date) - x_offset - bar_half_width
    xmax <- as.numeric(date) - x_offset + bar_half_width
    ymin <- previous_week_users
    ymax <- previous_week_users + number_of_users
    user_type = "New Users"
  } 
  
  result = data.frame(date = date,
                      user_type = user_type,
                      xmin = xmin,
                      xmax = xmax,
                      ymin = ymin,
                      ymax = ymax)
}

# Build out all of our rectangles. 
plot_data <- pmap_dfr(waterfall_df, calc_rect_coords)

# This trashes our nice date factor, so drop the one that came back
# and add on our original
plot_data <- select(plot_data, -date)
plot_data <- cbind(plot_data, waterfall_df$date)
plot_data <- plot_data %>% 
  select(date = `waterfall_df$date`,
         user_type,
         xmin, xmax, ymin, ymax)

# Display the head
head(plot_data, 20)

```

```{r waterfall_data_display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(plot_data,20))
```

# Data Visualization

With all that data prep, we can now plot the results. The way to read this is that, each week, there is a number of users. Then, for the next week, in order, we:

* Add the new users gained (new users for the week) -- green
* Then, subtract the (calculated) churned users -- red
* So we wind up with the total users for that week -- gray

The thing that may be a little misleading here is that we're making an assumption that "non-churned users" actually return to the site each subsequent week. But, that may not be too bad, especially if we're looking to illustrate that it's a constant battle to get new traffic to the site!

_As one note, ideally, this chart would also include horizontal lines connecting the top of each *Users* bar to the bottom of each *New Users* bar, from the top of each *New Users* bar to the top of each *Churned Users* bar, and from the bottom of each *Churned Users* bar to the top of each *Users* bar. But, that's an even bigger hassle, so it's not included here._

```{r plot_results, fig.width = 8}

# Define the theme
theme_waterfall <- theme_light() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.grid.major.x = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

# Build the actual plot. What we're actually doing here is drawing each bar as a
# separate rectangle.
gg_results <- ggplot(plot_data, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                                    fill = user_type, x=date)) +
  geom_rect() +
  scale_fill_manual(values = c("#f98e8e","#8ee26a","gray60")) +
  scale_y_continuous(label = comma, expand = c(0,0)) +
  labs(title = "Waterfall of Users Over Time", y="Users") +
  theme_waterfall

# Output the plot
gg_results

```


