---
title: "Pivoting Data (after Querying)"
---

This example pulls sessions by day for **device category** and **medium** and then displays them two different ways:

* As a heatmap with the totals for each device category / medium combination
* As a grid of sparklines for each device category / medium combination

. This is the advanced version of the [Pivoted Data (after Querying)](pivot-in-query.html) example.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

This is a simple query with just three dimensions and one metric. We're pulling the data at a daily level so we can plot the sparklines, but we're then going to "roll it up" for the heatmap totals.

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}

# Pull the data. See ?google_analytics_4() for additional parameters. The anti_sample = TRUE
# parameter will slow the query down a smidge and isn't strictly necessary, but it will
# ensure you do not get sampled data.
ga_data <- google_analytics_4(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = "sessions",
                              dimensions = c("date","medium","deviceCategory"),
                              anti_sample = TRUE)

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Munging

We've got two things we want to do here. Actually, one is a "want," and one is a "need:"

* **The Need** -- we need to aggregate/sum the sessions so we can do the heatmap
* **The Want** -- we want to convert the dimensions to factors with a "logical" order so that, more or less, both visualizations have sessions descending from the top left corner to the bottom right corner

Both of these are pretty straighforward to do.

```{r munging, message=FALSE, warning=FALSE, results='hide'}

# Remove the "date" component to get total sessions for the heatmap
ga_data_totals <- ga_data %>% 
  group_by(medium, deviceCategory) %>% 
  summarise(sessions = sum(sessions))

# Roll up to just be medium tables and then arrange the values for use in converting the
# medium column to a factor. This is one of those things that is pretty standard to need 
# to do. We're going to sort descending, but, for one of the data frames, we're actually
# going to need to reverse the order as we create the factors. This just comes with
# experimentation and experience.
medium_totals <- ga_data %>% 
  group_by(medium) %>% 
  summarise(sessions = sum(sessions)) %>% 
  arrange(-sessions)

# Do the same thing, but for deviceCategory. For this one, we actually want to sort 
# descending.
deviceCategory_totals <- ga_data %>% 
  group_by(deviceCategory) %>% 
  summarise(sessions = sum(sessions)) %>% 
  arrange(-sessions)

# Convert the medium and deviceCategory columns to factors in both of our main data frames.
ga_data$medium <- factor(ga_data$medium,
                         levels = medium_totals$medium)
ga_data$deviceCategory <- factor(ga_data$deviceCategory,
                         levels = deviceCategory_totals$deviceCategory)

ga_data_totals$medium <- factor(ga_data_totals$medium,
                         levels = rev(medium_totals$medium))    # Reversing the factor order
ga_data_totals$deviceCategory <- factor(ga_data_totals$deviceCategory,
                         levels = deviceCategory_totals$deviceCategory)

# Let's just check that the ga_data_totals we created looks pretty normal
head(ga_data_totals)

```

```{r munge-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data_totals))
```

# Data Visualization

We're going to break the standard for this site with this example by doing two visualizations:

* A heatmap of the data
* A similar grid, but with sparklines showing the data trended 

## The Heatmap

We'll make a little heatmap of the data in a pivoted fashion with ggplot. This requires two "geoms" -- `geom_tile()` to make the heatmap (the shaded grid), and then `geom_text()` to actually put the values in the heatmap. We'll use the `ga_data_totals` data frame for this.

The use of the `format()` function in the `label` argument is a handy little way to get commas displayed in numbers as the 000s separator (which means it's easy to swap out if you're in a locale where that is not the convention).

Note how the order of the rows and columns seems pretty logical -- from "biggest values" to "smallest values." This is entirely due to the data munging we did to convert **medium** and **deviceCategory** to factors and imposing some logic on the ordering of their levels.

There is a lot of additional adjustments that you may want to get into with this layout, but this site isn't intended to be a full tutorial on **ggplot2**, so we're going to call this "good enough."

```{r visualization, message=FALSE, warning=FALSE}

# Create the plot
gg <- ggplot(ga_data_totals, mapping=aes(x = deviceCategory, y = medium)) +
  geom_tile(aes(fill = sessions), colour = "grey30") +
  geom_text(aes(label = format(sessions, big.mark = ","))) +
  scale_fill_gradient(low = "white", high = "green") +    # Specify the gradient colors
  guides(fill = FALSE) +                                 # Remove the legend
  theme_light() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```

## A Grid of Sparklines

Now, we're going to use the _original_ data (albeit with **medium** and **deviceCategory** converted to factors) to plot a bunch of sparklines. Note that this is a fundamentally different technique than what was used for the heatmap. The heatmap used a "grid" geom -- `geom_text()` -- for the layout. For the sparklines, we're going to relyin on `facet_grid()`, which is not a geom, but, rather, specifies a grid into which individual `geom_line()` geoms will be displayed.

Once again, we've got relatively minimal tuning of the actual visualization here, and most of that is through adjustments through `theme()`. Obviously, this could be adjusted much more.

```{r visualization-sparklines, message=FALSE, warning=FALSE}

# Create the plot
gg_sparklines <- ggplot(ga_data, mapping=aes(x = date, y = sessions)) +
  geom_line() +
  facet_grid(medium ~ deviceCategory, switch = "y") +
  theme_light() +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "gray80"),
        panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black"),
        strip.text.y = element_text(angle = 0, hjust = 1),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg_sparklines

```


