---
title: "Pivoting Data within a Query"
---

**IMPORTANT NOTE: There is a bug in the Google Analytics v4 API that actually makes this example only partially functional (and not partially enough to make it work). Be sure to read the full intro here for an explanation!**

This example pulls sessions by **device category** and **medium**, but uses pivoting _within the call to Google Analytics_ so that the **device category** values are across the columns, while the **medium** values are in the rows. Generally, this is a bit of an odd thing to do for a couple of reasons:

* This may be personal preference, but it's awfully easy to do the pivoting with the `spread()` function in `dplyr` after it's been pulled as a flat/simple query.
* Generally, tidy data is, well, tidier. And, tidy data means pull the data...unpivoted.

But, it IS doable...except there is a bug in the API when it comes to pulling pivoted data -- a column can get dropped! Check out the [issue on Github](https://github.com/MarkEdmondson1234/googleAnalyticsR/issues/107) for an update.

Even without the bug, I consider this approach vastly inferior to the approach used in [Pivoting the Data (after Querying) example](pivot-after-query.html).

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

We have to, first, define what we're going to pivot. After that, it's a pretty straightforward query.

```{r data-pull, message=FALSE, warnings=FALSE, results='hide'}

# Start by defining the pivot object. See ?pivot_ga4() for details.
my_pivot_object <- pivot_ga4("deviceCategory",
                             metrics = "sessions")

# Pull the data. See ?google_analytics_4() for additional parameters. The anti_sample = TRUE
# parameter will slow the query down a smidge and isn't strictly necessary, but it will
# ensure you do not get sampled data.
ga_data <- google_analytics_4(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = "sessions",
                              dimensions = "medium",
                              pivots = list(my_pivot_object),
                              anti_sample = TRUE)

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warnings=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

This is the apparent/possible bug noted in the intro -- we seem to be missing a column! Check out the [issue on Github](https://github.com/MarkEdmondson1234/googleAnalyticsR/issues/107) for an update.

# Data Munging

The column headings are pretty ugly and redundant, and the **sessions** column isn't necessarily one we'd want to keep, so let's do a little cleanup.

```{r munging, message=FALSE, warnings=FALSE, results='hide'}

# Remove the 'sessions' column
ga_data <- select(ga_data, -sessions)

# Use a little regEx to strip out "deviceCategory" and "sessions" from the column names
names(ga_data) <- gsub("deviceCategory\\.(.*)\\.sessions","\\1", names(ga_data))

# Check out the result of our handiwork
head(ga_data)

```

```{r munge-display, echo=FALSE, message=FALSE, warnings=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Visualization

This could be a nice little heatmap, but to do that with ggplot2, we'd have to gather it up into a tidy format. Basically...unpivot it! That seems silly -- better to have pulled it unpivoted! I'm taking a moral stance and not jumping through that particular set of hoops for this example.