---
title: "Pivoting Data (after Querying)"
---

This example pulls sessions by **device category** and **medium** and then displays them in a pivoted fashion. This is the highly attractive cousin of the [pivot in the query itself](pivot-in-query.html) example.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

This is a simple query with just two dimensions and one metric. 

```{r data-pull, message=FALSE, warnings=FALSE, results='hide'}

# Pull the data. See ?google_analytics_4() for additional parameters. The anti_sample = TRUE
# parameter will slow the query down a smidge and isn't strictly necessary, but it will
# ensure you do not get sampled data.
ga_data <- google_analytics_4(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = "sessions",
                              dimensions = c("medium","deviceCategory"),
                              anti_sample = TRUE)

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warnings=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Munging

We actually don't need to do any data munging to prep for the data visualization, as the data is already in a nice, tidy format. But, if you want to pivot the data just to have it in a pivoted data frame, we can do that using the `spread()` function in `dplyr`.

```{r munging, message=FALSE, warnings=FALSE, results='hide'}

# Pivot the data
ga_data_pivoted <- ga_data %>% 
  spread(deviceCategory, sessions)

# Check out the result of our handiwork
head(ga_data_pivoted)

```

```{r munge-display, echo=FALSE, message=FALSE, warnings=FALSE}
# Nicer display of the head
kable(head(ga_data_pivoted))
```

# Data Visualization

Let's make a little heatmap of the data in a pivoted fashion with ggplot. This requires two "geoms" -- `geom_tile()` to make the heatmap (the shaded grid), and then `geom_text()` to actually put the values in the heatmap. **Note:** this uses the `ga_data` data frame that was pulled initially -- _not_ the `ga_data_pivoted` data frame that we created above. This is a subtle illustration of the elegance of the tidyverse, including `ggplot2`. If you appreciate that elegance, you are well on your way to R mastery.

The use of the `format()` function in the `label` argument is a handy little way to get commas displayed in numbers as the 000s separator (which means it's easy to swap out if you're in a locale where that is not the convention).

Note that there is not a logical/appropriate arrangement of the rows and columns, and the formatting is only minimally tweaked. This is one of the things addressed in the [intermediate-level verison of this example](pivot-sparklines-heatmap.html). 

```{r visualization, message=FALSE, warnings=FALSE}

# Create the plot
gg <- ggplot(ga_data, mapping=aes(x = deviceCategory, y = medium)) +
  geom_tile(aes(fill = sessions)) +
  geom_text(aes(label = format(sessions, big.mark = ","))) +
  scale_fill_gradient(low = "white", high = "green") +
  theme_light() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```
