---
title: "Simple Dynamic Segment - v3"
---

This example pulls the top 10 pages for the last thirty days, _for visits that occurred on a mobile device_. We'll doing this by defining a _dynamic segment_ using the v3 (older) version of the Google Analytics API. This returns the exact same results as these two examples, but through different means for defining/referencing the segment:

* **[Simple Dynamic Segment - v4](simple-dynamic-segment.html)** -- defining the segment using the v4 API syntax (which is super-cumbersome for a super simple segment like this one, but illustrative nonetheless)
* **[Apply Segment by Segment ID](segment-by-id.html)** -- simply using the ID of either a standard segment or a custom segment built in the Google Analytics web interface

All three approaches are perfectly acceptable.  

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

The trick to this is that we actually pass the [v3 dynamic segment definition](https://developers.google.com/analytics/devguides/reporting/core/v3/segments#dynamic) into the `segment_id` argument for `segment_ga4()`. It's not super-intuitive, because we're passing the _actual segment definition_ rather than an "id"... but when cramming v3 stuff into a v4 world, we've got to be a little forgiving, no?

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}

# For code readability, create a separate object with the segment definition.
mobile_segment_v3 <- "sessions::condition::ga:deviceCategory==Mobile"

# Create the segment object. See ?segment_ga4() for details. Note that, it doesn't
# really matter what we put for the name argument -- because this is pulling a v3
# dynamic segment, the name that appears in the output is just "Dynamic Segment."
my_segment <- segment_ga4("Mobile Sessions Only",
                          segment_id = mobile_segment_v3)

# Pull the data. See ?google_analytics_4() for additional parameters. Depending on what
# you're expecting back, you probably would want to use an "order" argument to get the
# results in descending order. But, we're keeping this example simple. Note, though, that
# we're still wrapping my_segment in a list() (of one element).
ga_data <- google_analytics_4(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = "pageviews",
                              dimensions = "pagePath",
                              segments = my_segment)

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)

```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Munging

Since we didn't sort the data when we queried it, let's go ahead and sort it here _and_ grab just the top 10 pages.

```{r munging, message=FALSE, warning=FALSE, results='hide'}

# Using dplyr, sort descending and then grab the top 10 values. We also need to make the
# page column a factor so that the order will be what we want when we chart the data.
# This is a nuisance, but you get used to it. That's what the mutate function is doing
ga_data_top_10 <- ga_data %>%
  arrange(-pageviews) %>% 
  top_n(10) %>% 
  mutate(pagePath = factor(pagePath,
                           levels = rev(pagePath)))

# Take a quick look at the result.
head(ga_data_top_10)
  
```

```{r munge-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data_top_10))
```

# Data Visualization

This won't be the prettiest bar chart, but let's make a horizontal bar chart with the data. Remember, in **ggplot2**, a horizontal bar chart is just a normal bar chart with `coord_flip()`.

```{r visualize, message=FALSE, warning=FALSE}

# Create the plot. Note the stat="identity"" (because the data is already aggregated) and
# the coord_flip(). And, I just can't stand it... added on the additional theme stuff to
# clean up the plot a bit more.
gg <- ggplot(ga_data_top_10, mapping = aes(x = pagePath, y = pageviews)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_light() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```
