---
title: "Medium Ordered by Sessions"
---

This example pulls the top 5 *mediums* by *sessions* for the last 30 days ordered in descending order by sessions.

# Setup/Config

```{r child='z_setup.Rmd'}
```

# Pull the Data

We just have to create an `order_type` object and then include that in the query.

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}

# Create an order type object.
order_sessions_desc <- order_type("sessions",
                                  sort_order = "DESCENDING",
                                  orderType = "VALUE")

# Pull the data. See ?google_analytics_4() for additional parameters. Note that we are
# NOT using anti_sample = TRUE. This is because we are setting max = 5, and the max
# argument has no impact when anti_sample = TRUE.
ga_data <- google_analytics(viewId = view_id,
                              date_range = c(start_date, end_date),
                              metrics = "sessions",
                              dimensions = "medium",
                              order = order_sessions_desc,
                              max = 5)

# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(ga_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(ga_data))
```

# Data Munging

Even though the data frame is ordered as we want it displayed, we still need to convert **medium** to a factor (with the levels in a reversed order) in order for the bar chart to be ordered from largest to smallest.

```{r data-munging, message=FALSE, warning=FALSE}
# Convert the medium to a factor so they will be ordered when plotted
ga_data$medium <- factor(ga_data$medium,
                         levels = rev(ga_data$medium))
```

# Data Visualization

This won't be the prettiest bar chart, but let's make a horizontal bar chart with the data. Remember, in **ggplot2**, a horizontal bar chart is just a normal bar chart with `coord_flip()`.

```{r visualize, message=FALSE, warning=FALSE}

# Create the plot. Note the stat="identity"" (because the data is already aggregated) and
# the coord_flip(). And, I just can't stand it... added on the additional theme stuff to
# clean up the plot a bit more.
gg <- ggplot(ga_data, mapping = aes(x = medium, y = sessions)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_light() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```